# Data augmentation

Some examples of adding extra columns to our play-by-play data in order to support particular analyses, or to make other data-wrangling tasks easier.

## The setter of a given attack

**Aim:** Identify the player who made the set associated with each attack (noting that some files might not have the setting action coded for all attacks, or even coded at all).

```{r eval = FALSE}
px <- px %>% mutate(set_player_id = case_when(skill == "Attack" & lag(skill) == "Set" & team == lag(team) ~ lag(player_id))
```

(The `set_player_id` column will be `NA` when there is no matching set coded before the attack.)

## Setter player IDs

**Aim:** Identify the `player_id` of the setter on court for each data row.

The play-by-play data frame has `home_player_id1`, `home_player_id2`, ..., `home_player_id6`, which give the `player_id` of the home team player currently in position 1, 2, ..., and 6. It also has a `home_setter_position` column, which tells us which position the home team setter is in (1--6). So to create a `home_setter_id` column we simply need to extract the value in the `home_player_idX` column for each row of the data frame, where `X` is the value in the `home_setter_position` at each row.

One way to do this is:

```{r eval = FALSE}
px <- px %>% mutate(home_setter_id = case_when(home_setter_position == 1 ~ home_player_id1,
                                               home_setter_position == 2 ~ home_player_id2,
                                               home_setter_position == 3 ~ home_player_id3,
                                               home_setter_position == 4 ~ home_player_id4,
                                               home_setter_position == 5 ~ home_player_id5,
                                               home_setter_position == 6 ~ home_player_id6))
```

But that is rather cumbersome and inelegant. A more concise method is:

```{r eval = FALSE}
px <- px %>% rowwise() %>% mutate(home_setter_id = cur_data()[[paste0("home_player_id", home_setter_position)]]) %>% ungroup
```

## Reception quality

**Aim:** Add a column that tells us the reception quality associated with each rally.

The reception quality for a given rally is found in the `evaluation` column of the row that has the `skill` value of "Reception". We just need to propagate this value to all other rows associated with that particular rally.

First we extract the reception quality for each rally (remember from the [Identifiers] section that the `point_id` value identifies the rally, but needs to be combined with `match_id` to be globally unique):

```{r receptionquality}
rq <- px %>% dplyr::filter(skill == "Reception") %>% group_by(match_id, point_id) %>%
  dplyr::summarize(reception_quality = if (n() == 1) .data$evaluation else NA_character_) %>% ungroup
```

So `rq` is a data.frame with the reception quality for each `match_id` and `point_id` combination. `reception_quality` will be `NA` if there was no reception (or more than one reception --- perhaps due to a scouting error?) involved in that point. Now join `rq` back to our full plays dataframe:

```{r receptionquality2}
px <- px %>% left_join(rq, by = c("match_id", "point_id"))
```

So now we have `reception_quality` for all rows of the data frame.

## Extracting file meta data

**Aim:** Add a column which has information from the file meta object.

It is possible to add parts of the meta data into your `px`. The data which you can add are as follows:

| meta\$match | meta\$more | meta\$video |
|-------------|------------|-------------|
| date        | referees   | file        |
| time        | spectators |             |
| season      | receipts   |             |
| league      | city       |             |
| phase       | arena      |             |
| home_away   | scout      |             |
| day_number  |            |             |
| mach_number |            |             |

: These steps follow the [Reading files] section steps with some additions to the `px` portion. Where `out` will end up becoming your `px` object / main data you will be using.

```{r eval = FALSE}
d <- dir("c:/data", pattern = "dvw$", full.names = TRUE)
lx <- lapply(d, dv_read, insert_technical_timeouts = FALSE)

px <- do.call(rbind, lapply(lx, function(dv) {
  out <- plays(dv)
  
  # For the lx[[1]]$meta$filename
  out$filename <- dv$meta$filename # basename(dv$meta$filename) is likely ideal
  
  # For the lx[[1]]$meta$match$date
  out$date <- dv$meta$match$date
  
  # for the meta$more$referees
  out$referees <- dv$more$referees

  # Return the data to px  
  return(out)
  }))
```
