# Data augmentation

Some examples of adding extra columns to our play-by-play data in order to support particular analyses, or to make other data-wrangling tasks easier.

## The setter of a given attack

**Aim:** Identify the player who made the set associated with each attack (noting that some files might not have the setting action coded for all attacks, or even coded at all).

```{r eval = FALSE}
px <- px %>% mutate(set_player_id = case_when(skill == "Attack" & lag(skill) == "Set" & team == lag(team) ~ lag(player_id))
```

(The `set_player_id` column will be `NA` when there is no matching set coded before the attack.)

## Setter player IDs

**Aim:** Identify the `player_id` of the setter on court for each data row.

The play-by-play data frame has `home_player_id1`, `home_player_id2`, ..., `home_player_id6`, which give the `player_id` of the home team player currently in position 1, 2, ..., and 6. It also has a `home_setter_position` column, which tells us which position the home team setter is in (1--6). So to create a `home_setter_id` column we simply need to extract the value in the `home_player_idX` column for each row of the data frame, where `X` is the value in the `home_setter_position` at each row.

One way to do this is:

```{r eval = FALSE}
px <- px %>% mutate(home_setter_id = case_when(home_setter_position == 1 ~ home_player_id1,
                                               home_setter_position == 2 ~ home_player_id2,
                                               home_setter_position == 3 ~ home_player_id3,
                                               home_setter_position == 4 ~ home_player_id4,
                                               home_setter_position == 5 ~ home_player_id5,
                                               home_setter_position == 6 ~ home_player_id6))
```

But that is rather cumbersome and inelegant. A more concise method is:

```{r eval = FALSE}
px <- px %>% rowwise() %>% mutate(home_setter_id = cur_data()[[paste0("home_player_id", home_setter_position)]]) %>% ungroup
```

## Reception quality

**Aim:** Add a column that tells us the reception quality associated with each rally.

The reception quality for a given rally is found in the `evaluation` column of the row that has the `skill` value of "Reception". We just need to propagate this value to all other rows associated with that particular rally.

First we extract the reception quality for each rally (remember from the [Identifiers] section that the `point_id` value identifies the rally, but needs to be combined with `match_id` to be globally unique):

```{r receptionquality}
rq <- px %>% dplyr::filter(skill == "Reception") %>% group_by(match_id, point_id) %>%
  dplyr::summarize(reception_quality = if (n() == 1) .data$evaluation else NA_character_) %>% ungroup
```

So `rq` is a data.frame with the reception quality for each `match_id` and `point_id` combination. `reception_quality` will be `NA` if there was no reception (or more than one reception --- perhaps due to a scouting error?) involved in that point. Now join `rq` back to our full plays dataframe:

```{r receptionquality2}
px <- px %>% left_join(rq, by = c("match_id", "point_id"))
```

So now we have `reception_quality` for all rows of the data frame.

## Extracting part of the file name

**Aim:** Add a column which has the file name or part of the file name

Dependent on your file naming convention it is possible to extract parts of the file name into your `px`. These steps do backtrack to the [Reading files] section.

```{r eval = FALSE}
d <- dir("c:/data", pattern = "dvw$", full.names = TRUE)
lx <- lapply(d, dv_read, insert_technical_timeouts = FALSE)
px <- do.call(rbind, lapply(lx, function(dv) {
  out <- plays(dv)
  out$filename <- dv$meta$filename
  out
  }))
```

Notice if you have your files nested in nested directories, this will return the entire file name path. To get the `basename` you can use the `basename` function.

```{r eval = FALSE}
d <- dir("c:/data", pattern = "dvw$", full.names = TRUE)
lx <- lapply(d, dv_read, insert_technical_timeouts = FALSE)
px <- do.call(rbind, lapply(lx, function(dv) {
  out <- plays(dv)
  out$filename <- basename(dv$meta$filename)
  out
  }))
```

If this is a dvw, return the `&` symbol and `.dvw` at the end.
